A Toy Analytic Transit Survey: Fixed Light Ratio Binaries

### 2017-08-30 memos/toy_analytic_survey_fixed_lightratio.tex
initial commit

\title{ A Toy Analytic Transit Survey }

@@

We derive the signal to noise distribution of detected planets in a toy transit 
survey. We give corresponding expressions for the number of detected planets. 
We then discuss the effect of various errors that one makes by ignoring 
binarity when deriving occurrence rates.

@@

Imagine the following transit survey:
\begin{itemize}
\item Take a magnitude-limited sample of ``stars'' of a single spectral class. 
For instance, G2V dwarfs.  In other words, choose all ``points'' on-sky with 
$m_\lambda < m_{\lambda,{\rm lim}}$, or equivalently with $F_\lambda > 
F_{\lambda,{\rm lim}}$, and also with colors that make you think they are this 
spectral class.
\item The true population of ``points'' (hereafter, systems) contains both 
single and double star systems. Single star systems have luminosity in the 
observed bandpass $L_1$, radii $R_1$, and effective temperature $T_{\rm eff,1}$.
Double star systems have luminosity in the observed bandpass $L_d = 
(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of the 
secondary to the primary. The ratio of the two number densities in a 
volume-limited sample is the binary fraction\footnote{The binary fraction is 
equivalent to the multiplicity fraction if there are no triple, quadruple, 
$\ldots$ systems.}.
\item The true population of planets around these stars is as follows:
	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.
    Any astrophysical difference in planet formation between singles and 
    binaries is captured by these two terms.
\end{itemize}

A signal to noise limited transit survey is then conducted. All planets with 
${\rm S/N} > {\rm (S/N)_{min}}$ are detected. The rest are not.
Note that while the S/N threshold does imply a minimum stellar flux about which 
planets of a given radius and period can be detected, this minimum flux is 
generally different from the minimum used to define the original magnitude cut 
for a sample [see Batalha et al, 2010].

@@

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C has lots 
of Keck time, and did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.

\item What about astronomer D, who additionally notes the importance of 
completeness?
\end{enumerate}

@@

In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
consequence the maximum distance to which binary stars will be selected is 
greater than that of single stars, simply as a consequence of imposing a 
magnitude cut.
The ratio of double to single systems is
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
&= {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{align}
In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$\sqrt{2}$ more binary systems than single systems in the sample.
Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
in single systems.

@@

@@

\section{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
These can be expressed individually. 
The former is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm S/N} > {\rm (S/N)_{min}}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, and $f_{s,({\rm S/N} > {\rm (S/N)_{min}})}$ is the 
fraction of these planets that have signal to noise greater than the minimum 
detection threshold.
\begin{equation}
N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,{\rm S/N} > {\rm (S/N)_{min}}},
\label{eq:N_det_d}
\end{equation}
where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
systems of the sample, and the completeness term must account for any 
differences in the signal to noise distribution that come from stellar 
binarity.

As a note of warning, Eqs.~\ref{eq:N_det_s} and~\ref{eq:N_det_d} ignore the 
geometric transit probability, $\approx R_\star / a$. This means we are 
assuming that every existing planet in this universe transits. This assumption 
can easily be rectified, if desired.

\subsection{Analytic completeness}
The only terms we have yet to compute are the completeness terms, $f_{i,{\rm 
S/N} > {\rm (S/N)_{min}}}$ for $i\in{\rm \{single,double\} }$. 
We proceed as follows.

The interesting term in Eq.~\ref{eq:snr_ivory_tower} that changes between stars 
of the same binarity class in our idealized sample is the square root of 
$F^{\rm N}_\gamma$.
This is the term that leads to a distribution of signal to noises for different 
stars.
The completenesses $f_{i,{\rm S/N} > {\rm (S/N)_{min}}}$ can be directly 
expressed in terms of those probability density functions:
\begin{equation}
f_{i,{\rm S/N} > {\rm (S/N)_{min}}} = 
	\int_{{\rm (S/N)_{min}}}^{\infty} 
		{\rm d}\left(\frac{{\rm S}}{{\rm N}}\right)_i \ 
		{\rm prob}\left(\frac{{\rm S}}{{\rm N}}\right)_i.
\label{eq:completeness_long}
\end{equation}
We keep the subscript $i$ because the signal to noise distributions are 
different for the cases of single star systems ($i=s$) and double star systems 
($i=d$).
Notably:
\begin{itemize}
	\item The dilution differs (Eq.~\ref{eq:dilution}).
	\item The photon number flux from the system differs.
\end{itemize}

To simplify notation, we let $x_i \equiv {\rm (S/N)}_i$, and rewrite 
Eq.~\ref{eq:completeness_long} as
\begin{equation}
f_{i,x > x_{\rm min}} = 
\int_{{\rm x_{\rm min}}}^{\infty} 
{\rm d}x_i \ 
{\rm prob}(x_i).
\label{eq:completeness_short}
\end{equation}

\subsection{Deriving ${\rm prob}(x_i)$}
We want expressions for the probability density function of the observed signal 
to noise ratio, ${\rm prob}(x_i)$, for both the single and binary 
system case.

First, note that a star placed uniformly in the volume of the search space will 
have a probability density function for its distance $r$ from the origin of
\begin{equation}
{\rm prob} (r) = \frac{3 r^2}{d_{\rm max}^3},
\end{equation}
where the appropriate maximum distances should be substituted per 
Eq.~\ref{eq:d_max}.
Noting the transformation rule for probability density functions, we can 
evaluate the probability of a star having a observed number flux $F^{\rm 
N}_{i,\gamma}$ in the bandpass,
\begin{align}
{\rm prob} (F^{\rm N}_{i,\gamma}) 
&= {\rm prob}(r(F^{\rm N}_{i,\gamma}))
	\left| \frac{{\rm d} r}{{\rm d} F^{\rm N}_{i,\gamma}} \right| \\
&= \frac{3}{2 d_{\rm max}^3} c_i^{3/2} (F^{\rm N}_{i,\gamma})^{-5/2},
\label{eq:pdf_observed_flux}
\end{align}
where in the latter equality we have written a ``number luminosity'' $c_i$ 
(units of inverse time) defined for $i\in{\rm \{single,double\} }$ as
\begin{equation}
c_i = 
\Bigg\{\begin{array}{lr}
R_1^2 F^{\rm N}_{s1,\gamma}, & {\rm\ if\ single}\\
R_1^2 F^{\rm N}_{s1,\gamma} + R_2^2 F^{\rm N}_{s2,\gamma} & {\rm\ if\ double}.
\end{array}
\label{eq:c_i}
\end{equation}
In Eq.~\ref{eq:c_i}, $F^{\rm N}_{s1,\gamma}$ and $F^{\rm N}_{s2,\gamma}$ are 
the photon number fluxes at the surfaces of the stars. To derive 
Eq.~\ref{eq:pdf_observed_flux}, we simply scaled these by the distance:
\begin{equation}
F^{\rm N}_{i,\gamma} = \frac{c_i}{r^2}.
\label{eq:flux_at_detector}
\end{equation}

The surface photon number fluxes $F^{\rm N}_{si,\gamma}$ in Eq.~\ref{eq:c_i} 
are usually evaluated numerically, by convolving the wavelength-specific photon 
flux density of a star with the dimensionless spectral response function of the 
instrument. In other words,
\begin{equation}
F^{\rm N}_{s,\gamma} = \int F_\lambda T_\lambda \, {\rm d}\lambda.
\label{eq:pfd}
\end{equation}
The wavelength-specific photon flux density $F_\lambda$ [${\rm 
ph\,cm^{-2}\,s^{-1}\,\AA^{-1}}$] could be from Pickles' library, or could be 
a blackbody function. The transmission function is, up to factor of order 
unity, a step function between two wavelengths $\lambda_{\rm min}$ and 
$\lambda_{\rm max}$.
If we assume a blackbody source, Eq.~\ref{eq:pfd} becomes
\begin{equation}
F^{\rm N}_{s,\gamma} = 8\pi c \left(\frac{k T}{h c}\right)^3 
\int_{hc/(\lambda_{\rm max} kT)}^{hc/(\lambda_{\rm min} kT)}
\frac{u^2}{e^u - 1} \,{\rm d} u,
\end{equation}
which can be evaluated numerically\footnote{It may help in the numerics to note 
that infinite series representations of this type of dimensionless integral 
exist and converge quickly. For instance, one can show that
\begin{equation}
\int_{0}^{a} \frac{u^3}{e^u -1} \,{\rm d}u = \sum_{n=1}^{\infty} \frac{6}{n^4} 
- \frac{e^{-an}}{n^4} (6 + 6an + 3(an)^2 + (an)^3).
\nonumber
\end{equation}
A similar expression exists for the similar integral in the text. 
[Michels 1968] explains an analogous derivation.
}.

The importance of the functional form of $F^{\rm N}_{s,\gamma}$ is that it is 
to first order only a function of the blackbody temperature and the bandpass 
wavelength bounds.
Thus in the most general case $c_i(R_1, R_2, T_1, T_2, \lambda_{\rm min}, 
\lambda_{\rm max})$ and nothing else.
The only random variable involved in the flux being received at the detector is 
$r$, so we can indeed write the flux received at the detector as in 
Eq.~\ref{eq:flux_at_detector}.

We can finally write out the probability density functions for the signal to 
noise ratios in the single and double-star cases by using the transformation 
rule for pdfs, and applying Eq.~\ref{eq:snr_ivory_tower}.
For single stars,
\begin{equation}
{\rm prob}(x_s) = \frac{3}{d_{\rm max, s}^3} c_s^{3/2} \delta^3 \left( A T_{\rm 
dur} N_{\rm tra}\right)^{3/2} x_s^{-4}.
\label{eq:prob_xs}
\end{equation}
Analogously for double stars,
\begin{equation}
{\rm prob}(x_d) = \frac{3}{d_{\rm max, d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 
\left( A T_{\rm dur} N_{\rm tra}\right)^{3/2} x_d^{-4}.
\label{eq:prob_xd}
\end{equation}

\subsection{Number of detected planets}

Performing the integrals of Eq.~\ref{eq:completeness_short}, we get:

\begin{align}
N_{{\rm det},s} &= N_s \Gamma_{t,s} f_{s,{\rm S/N} > {\rm (S/N)_{min}}} \\
&= N_s \Gamma_{t,s} \frac{1}{d_{\rm max,s}^3} c_s^{3/2} \delta^3 \left(A 
T_{\rm dur} N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3},
\end{align}
and
\begin{align}
N_{{\rm det},d} &= 2 N_d \Gamma_{t,d} f_{d,{\rm S/N} > {\rm (S/N)_{min}}} \\
&= 2 N_d \Gamma_{t,d} 
\frac{1}{d_{\rm max,d}^3} c_d^{3/2} (\mathcal{D}\delta)^3 \left(A T_{\rm dur} 
N_{\rm tra}\right)^{3/2} x_{\rm min}^{-3}.
\end{align}

Formally, the $f_{s,{\rm S/N} > {\rm (S/N)_{min}}}$ term should really be 
written as ${\rm min}(1, \ldots)$, where $(\ldots)$ is the given expression. 
This ensures that the fraction of planets above the signal to noise threshold 
is less than 1.

The number of detected planets $N_{\rm det}$ is the sum of the two previously 
written equations, and can be written
\begin{align}
N_{\rm det} = 
&\left( \frac{\delta}{x_{\rm min}} \right)^3 (A T_{\rm dur} N_{\rm tra})^{3/2} 
\nonumber\\
&\quad \times
\left[ N_s \Gamma_{t,s} \frac{c_s^{3/2}}{d_{\rm max,s}^3}  +
	   2 N_d \Gamma_{t,d} \frac{c_d^{3/2}}{d_{\rm max,d}^3}  \right].
\label{eq:N_det}
\end{align}
All terms can be input to a computer, and checked against a Monte Carlo 
simulation if desired.

@@

The {\it total} occurrence rate (number of detections divided number of 
``stars'') for Astronomer A would be $N_{\rm det}/(N_s + N_d)$. However, even 
though Astronomer A does not know about binaries, the radii he derives for any 
planets in binary systems are too small, by a factor $\sqrt{\mathcal{D}}$.
The question is what occurrence rate is derived for planets of radius $R_p$ and 
period $P$.
The answer is thus
\begin{equation}
\Gamma_{\rm A,\ planets\ of\ R_p} = \frac{N_{\rm det,s}}{N_s + N_d}.
\end{equation}

This astronomer will also think there is a second population of planets, with 
radius $R_p \sqrt{\mathcal{D}}$, and will thus rush to Nature claiming to also
have derived a second occurrence rate,
\begin{equation}
\Gamma_{\rm A,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}}{N_s 
+ N_d}.
\end{equation}

@@

By the same token as above,
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p} = \frac{N_{\rm det,s}}{N_s + 2N_d},
\end{equation}
and
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm det,d}}{N_s 
	+ 2N_d}.
\end{equation}

\section{Astronomer C counts host stars correctly and figures out diluted radii}
Astronomer C did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.
Thus, she also knows which planets are in binary systems, and which are in 
single star systems.

She knows that the purported population of planets with radii $R_p 
\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
radii $R_p$. She computes an occurrence rate
\begin{align}
\Gamma_{\rm C,\ planets\ of\ R_p} &= 
\frac{N_{\rm det,s} + N_{\rm det,d}}{N_s + 2N_d} \\
&= \frac{N_{\rm det}}{N_s + 2N_d},
\end{align}

the closest yet to the true rate (Sec.~\ref{sec:true_rate}).

\section{Astronomer D counts host stars correctly, figures out diluted radii, 
and accounts for completeness}

Astronomer D, just like Astronomer C, knows which detections were around 
binaries and the associated radius correction.

Astronomer D notes that the number of detected planets is not equal to the 
actual number of planets in the sample.
They must account for completeness.
They do injection recovery, and derive independent estimates for their 
completeness functions about single stars, $f_{s,x>x_{\rm min}} \equiv f_s$ and 
about double star systems, $f_{d,x>x_{\rm min}} \equiv f_d$.
With this knowledge in hand, they compute the respective single and binary 
occurrence rates
\begin{equation}
\Gamma_{t,s} = \frac{N_{\rm det,s}}{N_s f_s},
\end{equation}
\begin{equation}
\Gamma_{t,d} = \frac{N_{\rm det,d}}{2 N_d f_d}.
\end{equation}

With these in hand, they derive the overall occurrence rate
\begin{align}
\Gamma_{\rm D,\ planets\ of\ R_p} 
&= \frac{N_{\rm det,s}/f_s + N_{\rm det,d}/f_d  }{N_s + 2 N_d} \\
&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
\end{align}

Although we admit it's been a bit of a slog, it happens that Astronomer D's 
occurrence rate is the true occurrence rate (cf. Eq.~\ref{eq:true_occ}).

All it takes is $\approx$ a full semester at Keck, a system-by-system analysis, 
and perfect understanding of the completeness of the detection efficiency for 
single and double star systems.

@@

As a check on the preceding analytic development, we implemented a Monte Carlo 
simulation of this idealized transit survey.
To run the survey, we defined the instrument specifications (detector area and 
transmission function), the stellar population (binary fraction, total number 
density of a given stellar class, binary light ratio, fixed stellar 
properties), the planet population (fixed planet radius, period, and occurrence 
rate about single and binary stars), and finally the survey parameters 
(observing baseline, minimum SNR for ``detection'').
We then randomly drew star positions, randomly assigned planets to stars in 
single and binary systems, and computed the resulting signal to noise 
(Eq.~\ref{eq:snr_ivory_tower}) with which 
the transits would be observed.
As in the preceding analytics\footnote{TODO could be to generalize away from 
this}, we assumed ``twin'' binaries (same stellar radii, same effective 
temperature, and dilution does not depend on which stellar binary is the 
``target'').

The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
indicate that the analytic probability distribution functions 
Eqs.~\ref{eq:prob_xs},~\ref{eq:prob_xd}, and the number of 
detections (Eq.~\ref{eq:N_det}) are correct.

A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
distance $r$) the SNR distribution for planets in binaries is poorer than that 
of planets in single star systems.
We can see analytically that this can be written only as function of the 
binary light ratio:
\begin{equation}
\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1}.
\end{equation}
Deriving this simple form requires noting that the ratios of the 
bandpass-specific number luminosities (Eq.~\ref{eq:c_i}) is equal to the ratio 
of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
must be included).

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
	\end{center}
	\caption{Comparison of analytic and numeric probability 
	density functions of the SNR in an idealized transit survey.
	The analytic lines are Eqs.~\ref{eq:prob_xs} and~\ref{eq:prob_xd} for the 
	planet populations orbiting single and binary stars. The underlying stepped 
	histogram is output from Monte Carlo simulations. Poisson noise leads to 
	a small deviation at the faint and bright limits, but the numerics and 
	analytics otherwise agree.
		 }
	\label{fig:snr_dist_analytic_v_numeric}
\end{figure}

@@

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
A,R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.

>>>>

\title{ A Toy Analytic Transit Survey: Fixed Light Ratio Binaries }

@@

We discuss various errors one can make by ignoring binarity when deriving 
occurrence rates.
In this memo, we assume the stars in any binary system have fixed light ratio.

@@

Imagine the following survey, similar to that outlined by [Pepper et al, 2003]:
\begin{enumerate}
\item You are going to observe the entire sky for a duration $T_{\rm obs}$, 
with a detector of area $A$, and known bandpass. Your detector is photon-noise 
limited.
%
\item You are interested only in detecting planets of radius $R_p$, and orbital 
period $P$. For instance, $R_p=R_\oplus$, $P={\rm 1\ year}$.
%
\item You only want to detect them around stars of radius $R_1$, and luminosity 
$L_1$. For instance, G2V dwarfs.
%
\item Since your detections will be S/N limited, you want only to observe stars 
for which your target can be detected with ${\rm S/N} > {\rm (S/N)_{min}}$.
%
\item For a photon-noise limited survey, the signal to noise limit is 
equivalent to a magnitude (flux) limit. So you select all the points on the sky 
above a flux limit, \textit{i.e.} with flux $F_{\rm pt} > F_{\rm min}$, for 
$F_{\rm min}$ defined by your target planet and stellar types, and your survey 
specifications.
%
\item You carry out a transit survey, and detect many transiting planets.
\end{enumerate}

You now wish to derive an occurrence rate for planets of radius $R_p$ and 
orbital period $P$.
Assume your universe is a universe in which:
\begin{itemize}
\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises both single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
  In this memo, $\gamma_R$ is assumed to be a constant across the population of
  star systems.
  The ratio of the two number densities in a 
	volume-limited sample is the binary fraction\footnote{The binary fraction 
	is equivalent to the multiplicity fraction if there are no triple, 
	quadruple, $\ldots$ systems.}.
\item The true population of planets around these stars is as follows:
	\subitem A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have a planet of radius $R_p$, with orbital period $P$.
	\subitem A fraction $\Gamma_{t,d}$ of each star in a double star 
	system has a planet of radius $R_p$, with orbital period $P$. For instance, 
	if $\Gamma_{t,s} = \Gamma_{t,d} = 0.1$, on average each double 
	system contributes 0.2 planets, and each single system 0.1 planets.
    Any astrophysical difference in planet formation between singles and 
    binaries is captured by these two terms.\footnote{We'll see that it's 
    an extreme limit to assume the primary and secondary of binary systems host 
    planets at equal rates if the secondary is non-identical to the primary.}
\end{itemize}

@@

\item What about astronomer C, who accounts for ``2 for 1'' \textit{and} 
misclassification due to diluted radii? In other words, astronomer C did a 
combination of high resolution imaging and RV followup on every candidate, and
correctly classifies the planetary radii in every case.
\end{enumerate}

@@

In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$, and as a 
consequence the maximum distance to which binary stars will be selected is 
greater than that of single stars, simply as a consequence of imposing a 
magnitude cut.
The ratio of double to single systems, assuming $\gamma_R$ is a constant for 
the entire population, is
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s} \left(\frac{d_{{\rm max}, d} }{d_{{\rm max}, s}}\right)^3 \\
&= {\rm BF} \times (1+\gamma_R)^{3/2}.
\end{align}
In the nominal case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$\sqrt{2}$ more binary systems than single systems in the sample.
Correspondingly, there are $2\sqrt{2}$ more stars in binary systems than stars 
in single systems.

@@

\subsection{Actually finding $F_{\rm lim}$}

To express $F_{\rm lim}$ in terms of the survey's defining parameters, we need 
to convert between ${\rm (S/N)_{lim}}$ and $F_{\rm lim}$. We proceed as follows.

Using Eq.~\ref{eq:snr_ivory_tower}, we can write the minimum number flux of 
photons required for a detection at threshold as
\begin{equation}
F_{\rm lim}^N = \left[ \left({\rm \frac{S}{N}}\right)_{\rm min} \frac{1}{\delta 
\mathcal{D}} \right]^2 \frac{1}{A N_{\rm tra} T_{\rm dur}}.
\end{equation}
To convert this to $F_{\rm lim}$, multiply by the average photon energy in the 
bandpass.

@@

\section{How many planets are detected?}
The total number of planet detections is the sum of the number of planets 
detected in single star systems $N_{{\rm det},s}$ and the number of planets 
detected in double star systems $N_{{\rm det},d}$.
These can be expressed individually. 
Since we selected stars for which there was enough light to make detections, 
there is no need to compute the ${\rm S/N}$ distribution of 
``threshold-crossing events''.

The number of planets detected in single star systems is
\begin{equation}
N_{{\rm det},s} = N_s \Gamma_{t,s} f_{s,{\rm geom}},
\label{eq:N_det_s}
\end{equation}
where the product $N_s \Gamma_{t,s}$ is the number of planets in the single 
star systems of the sample, and $f_{s,{\rm geom}}$ is the geometric transit 
probability.
The number of planets detected in double star systems is
\begin{equation}
N_{{\rm det},d} = 2 N_d \Gamma_{t,d} f_{d,{\rm geom}},
\label{eq:N_det_d}
\end{equation}
where now $2 N_d \Gamma_{t,d}$ is the number of planets in the double star 
systems of the sample.
The number of detected planets $N_{\rm det}$ is the sum of the two previously 
written equations.

@@

The {\it total} occurrence rate (number of planets divided number of 
``stars'') for Astronomer A would be $(N_{\rm det}/f_{s,{\rm geom}})/(N_s + 
N_d)$. However, even though Astronomer A does not know about binaries, the 
radii he derives for any planets in binary systems are too small, by a factor 
$\sqrt{\mathcal{D}}$. The question asks what occurrence rate is derived for 
planets of radius $R_p$ and period $P$.
The answer is
\begin{align}
\Gamma_{\rm A,\ planets\ of\ R_p} &= \frac{N_{\rm det,s}/f_{s,{\rm geom}}}{N_s 
+ N_d}
\label{eq:A_0} \\
&= \frac{\Gamma_{t,s} N_s}{N_s + N_d}
\end{align}

This astronomer will also think there is a second population of planets, with 
radius $R_p \sqrt{\mathcal{D}}$ (a constant number for $\gamma_R = {\rm 
constant}$). He will then also claim
have derived a second occurrence rate,
\begin{equation}
\Gamma_{\rm A,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm 
det,d}/f_{d,{\rm geom}}}{N_s + N_d},
\end{equation}

where at least for the twin binary case the geometric completeness term is the 
same as for Eq.~\ref{eq:A_0}.

@@

By the same token as above,
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p} = \frac{N_{\rm det,s}/f_{s,{\rm geom}}}{N_s + 
2N_d},
\end{equation}
and
\begin{equation}
\Gamma_{\rm B,\ planets\ of\ R_p \sqrt{\mathcal{D}}} = \frac{N_{\rm 
det,d}/f_{d,{\rm geom}}}{N_s 
	+ 2N_d}.
\end{equation}

\section{Astronomer C counts host stars correctly and figures out diluted radii}
Astronomer C did high resolution imaging followup on every candidate, and 
correctly classifies the planetary radii.
Thus, she also knows which planets are in binary systems, and which are in 
single star systems.
Since there are no completeness corrections, she has everything she needs.

She knows that the purported population of planets with radii $R_p 
\sqrt{\mathcal{D}}$ does not exist. All detected planets from this survey have 
radii $R_p$. She computes an occurrence rate
\begin{align}
\Gamma_{\rm C,\ planets\ of\ R_p} &= 
\frac{N_{\rm det,s}/f_{s,{\rm geom}} + N_{\rm det,d}/f_{d,{\rm geom}}}{N_s + 
2N_d} \\
&= \frac{\Gamma_{t,s} N_s + 2\Gamma_{t,d} N_d}{N_s + 2 N_d}.
\end{align}

With $\approx$ a full semester at Keck, a system-by-system analysis, 
and perfect understanding of the completeness of the detection efficiency for 
single and double star systems, Astronomer C has found the true occurrence rate 
(cf. Eq.~\ref{eq:true_occ}).

@@

As a check on the preceding analytic discussion, we implemented a Monte Carlo 
simulation of this idealized transit survey.
To run the survey, we defined the instrument specifications (detector area and 
transmission function), the stellar population (binary fraction, total number 
density of a given stellar class, binary light ratio, fixed stellar 
properties), the planet population (fixed planet radius, period, and occurrence 
rate about single and binary stars), and finally the survey parameters 
(observing baseline, minimum SNR for ``detection'').
We then randomly drew star positions, randomly assigned planets to stars in 
single and binary systems, and computed the resulting signal to noise 
(Eq.~\ref{eq:snr_ivory_tower}) with which 
the transits would be observed.
As in the preceding analytics, we assumed ``twin'' binaries (same stellar 
radii, same effective temperature, and dilution does not depend on which 
stellar binary is the ``target'').

The results are shown in Fig.~\ref{fig:snr_dist_analytic_v_numeric}, and 
indicate that the analytic probability distribution functions derived in an 
earlier version of this memo are correct.

A point evident in Fig.~\ref{fig:snr_dist_analytic_v_numeric} is that, for 
fixed planet parameters, and fixed stellar parameters ($R_\star, L_\star$, and 
distance $r$), and a constant $\gamma_R$ across the population, the SNR 
distribution for planets in binaries is poorer than that of planets in single 
star systems.
We can see analytically that this can be written only as function of the 
binary light ratio:
\begin{equation}
\frac{{\rm prob}(x_d)}{{\rm prob}(x_s)} =(1 + \gamma_R)^{-1}.
\end{equation}
Deriving this simple form requires noting that the ratios of the 
bandpass-specific number luminosities is equal to the ratio 
of the bandpass-specific energy luminosities (otherwise a term with $c_s/c_d$ 
must be included).

\begin{figure}[!t]
	\begin{center}
		\includegraphics[scale=0.4]{figures/snr_distribution.pdf}
	\end{center}
	\caption{Comparison of analytic and numeric probability 
	density functions of the SNR in an idealized transit survey.
	The analytic lines are from a previous version of this memo 
	(\texttt{20170728}) for the planet populations orbiting single and binary 
	stars. The underlying stepped histogram is output from Monte Carlo 
	simulations. Poisson noise leads to a small deviation at the faint and 
	bright limits, but the numerics and analytics otherwise agree.
		 }
	\label{fig:snr_dist_analytic_v_numeric}
\end{figure}

@@

For the nominal G2V dwarf case of ${\rm BF}=0.45$ with twin binaries ($\gamma_R 
= 1$) and $\Gamma_{t,d} = \Gamma_{t,s}$ this gives a relative error of $127\%$.
For instance, in the numerical simulation corresponding to 
Fig.~\ref{fig:snr_dist_analytic_v_numeric}, Astronomer A finds $\Gamma_{\rm 
A,R_p} = 0.22$, while Astronomer 
D derives the true (input) occurrence rate of $\Gamma_{\rm D,R_p} = 0.5$.
For the more ``realistic'' case of $\gamma_R = 0.1$, $\Gamma_{t,d} = 
\Gamma_{t,s}$, the relative error is $52\%$. Not as bad, and this latter number 
should be looked at skeptically because the ``equal occurrence rates for 
primary and secondary'' assumption may be unlikely with this light ratio as a 
median.
